OpenMPIR

### 2017-10-06 paper.tex
Fixed citation for Barcelona benchmarks.
edits to memory model discussion

To evaluate the use of Tapir we compared performance on the Barcelona OpenMP Task 
Suite to several existing OpenMP implementations. The suite is a set of tests
intended to test performance of OpenMP implementations using both irregular and
regular tasking. The following list provides abridged descriptions of each
benchmark taken verbatim from the test suite distribution \cite{bots}.

@@

The surprising fact that program behavior wasn't changed significantly is worth
discussing. We surmise that it is likely due to the semantics of variables
borrowed from the Cilk codegen and their relation to the OpenMP specification
of semantics. In particular, \emph{OpenMP specifies a very relaxed memory model
for shared variables}. This results in the Cilk implementation of only writing
to a shared variable at the end of a parallel section being a perfectly valid
implementation of OpenMP's specification. In contrast, existing OpenMP
implementations often write to shared variables on every write. This is a
potential for performance discrepancy because of differing
implementation-defined behavior and we will return to it when discussing
performance on the Barcelona OpenMP task suite. There is of course the fact
that technically, by copying the value of variables back to the surrounding
context even for \texttt{private} variables, one isn't following the
specification. We surmise that in practice this hasn't had an effect because
\texttt{private} variables are generally used for performance rather than their
semantic properties.

@@

This work was supported by the Director, Office of Advanced Scientific
Computing Research, Office of Science, of the United States Department
of Energy, under the guidance of Dr. Sonia Sachs. This research was
supported by US National Science Foundation under CRII ACI 1565338,
CNS 1527076, and CNS 1217948 and by the Exascale Computing Project
(17-SC-20-SC), a collaborative effort of the U.S. Department of Energy
Office of Science and the National Nuclear Security Administration.
Los Alamos National Laboratory is managed and operated by Los Alamos
National Security, LLC (LANS), under contract number DE-AC52-06NA25396
for the Department of Energy's National Nuclear Security
Administration (NNSA). Sandia National Laboratories is a multimission
laboratory managed and operated by National Technology and Engineering
Solutions of Sandia, LLC., a wholly owned subsidiary of Honeywell
International, Inc., for the U.S. Department of Energy's National
Nuclear Security Administration under contract DE-NA-0003525.

>>>>

To evaluate the use of Tapir we compared performance on the Barcelona OpenMP Task 
Suite to several existing OpenMP implementations. The suite is a set of tests
intended to test performance of OpenMP implementations using both irregular and
regular tasking. The following list provides abridged descriptions of each
benchmark taken verbatim from the test suite distribution \cite{barcelona}.

@@

The surprising fact that program behavior wasn't changed significantly is worth
discussing. We surmise that it is likely due to the semantics of variables
borrowed from the Cilk codegen and their relation to the OpenMP specification
of semantics. In particular, \emph{OpenMP specifies a relaxed memory model
for shared variables}, in which writes by one thread need only be visible at
to other threads at explicit or implicit flushes. 
This results in the Cilk implementation of only writing
to a shared variable at the end of a parallel section being a valid
implementation of OpenMP's specification. In contrast, existing OpenMP
implementations sometimes write shared variables to memory on every write access. This is a
potential for performance discrepancy because of differing 
implementation behavior.
%Stephen:This topic is not discuessed later, but was touched on before -- moved?
%and we will return to it when discussing
%performance on the Barcelona OpenMP task suite. 
There is of course the fact
that technically, by copying the value of variables back to the surrounding
context even for \texttt{private} variables, one isn't following the
specification. We surmise that in practice this hasn't had an effect because
\texttt{private} variables are often used for input parameters in the 
progams or for performance rather than their semantic properties.
Also, \texttt{firstprivate} is the default data-sharing attribute for tasks.

@@

This work was supported by the Director, Office of Advanced Scientific
Computing Research, Office of Science, of the United States Department
of Energy, under the guidance of Dr. Sonia Sachs. This research was
supported by US National Science Foundation under CRII ACI 1565338,
CNS 1527076, and CNS 1217948 and by the Exascale Computing Project
(17-SC-20-SC), a collaborative effort of the U.S. Department of Energy
Office of Science and the National Nuclear Security Administration.
Los Alamos National Laboratory is managed and operated by Los Alamos
National Security, LLC (LANS), under contract number DE-AC52-06NA25396
for the Department of Energy's National Nuclear Security
Administration (NNSA).

Sandia National Laboratories is a multimission laboratory managed and
operated by National Technology and Engineering Solutions of Sandia,
LLC., a wholly owned subsidiary of Honeywell International, Inc., for
the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-NA-0003525.
